---
title: "EDAV_PSet5A"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)  
```

## R Markdown

```{r cars}

library(dplyr)
library(ggplot2)
library(tidyverse)
library(maps)
library(GGally)
library(RColorBrewer)
library(parcoords)

```
The dataset has 4 tables corresponding to inter-state wars, intra-state wars, non state wars and extra state wars. 
Many wars transformed to other wars that took place later on. We found that 12 civil intra state wars transformed to inter state wars, 3 inter state wars transformed to extra state wars and 2 intra state wars transformed to extra state wars. We were able to find these transformed wars and store them as shown below:



```{r}
inter_state=read.csv('/home/hima/Downloads/data_raw/Inner-state Wars/Inter-StateWarData_v4.0.csv')
intra_state=read.csv('/home/hima/Downloads/data_raw/Intra-state Wars/Intra-StateWarData_v4.1.csv')
Extra_state=read.csv('/home/hima/Downloads/data_raw/Extra-state Wars/Extra-StateWarData_v4.0.csv')
military_expenditure_main = read.csv('/home/hima/Downloads/military-expenditure-as-a-share-of-gdp (1).csv')
colnames(military_expenditure_main)[4] <- "MilitaryExpenditure"
primary_school_main = read.csv('/home/hima/Downloads/primary-enrollment-selected-countries.csv')
colnames(primary_school_main)[4] <- "SchoolEnrolment"
HDI_main = read.csv('/home/hima/Downloads/human-development-index-escosura.csv')
colnames(HDI_main)[4] <- "HDI"
```

<!-- Intra state wars are further categorised into 3 types as follows: -->
<!-- ```{r} -->
<!-- Civil_intra_state=intra_state%>%filter(WarType==4|WarType==5) -->
<!-- Regional_internal_intra_state=intra_state%>%filter(WarType==6) -->
<!-- Intercommunal_intra_state=intra_state%>%filter(WarType==7) -->
<!-- ``` -->


World War 1

```{r}
#world_war1:: Years: 1914-1918
world_war1=inter_state%>%filter(WarName=='World War I')
world_war1$StateName=as.character(world_war1$StateName)
```


```{r}
#maps
world <- map_data("world")
world_war1 <- world_war1 %>%
  ## Recode certain entries
  mutate(StateName = recode(str_trim(StateName), "United States of America" = "USA",
                            "United Kingdom" = "UK"))
```

```{r}
world_war1['start_date'] <- as.Date(with(world_war1, paste(StartYear1, StartMonth1, StartDay1,sep="-")), "%Y-%m-%d")
world_war1['end_date'] <- as.Date(with(world_war1, paste(EndYear1, EndMonth1, EndDay1,sep="-")), "%Y-%m-%d")
world_war1['number_of_days']=world_war1['end_date']-world_war1['start_date']
```



Military Expenditure Trends

```{r}
library(ggplot2)
military_expenditure <- subset(military_expenditure_main, Entity %in% world_war1$StateName)
  ggplot(military_expenditure,aes(Year, y = MilitaryExpenditure, colour = Entity, group = Entity)) +
  geom_line() +
  labs(title = "Trends in Military Expenditures of countries in WW1", y = "Expenditure", x = "")
  
  library(ggplot2)
military_expenditure <- subset(military_expenditure_main, Entity %in% world_war1$StateName)
military_expenditure <- subset(military_expenditure, Year >= min(world_war1$StartYear1)-10)
military_expenditure <- subset(military_expenditure, Year <= max(world_war1$EndYear1)+10)
 ggplot(military_expenditure,aes(Year, y = MilitaryExpenditure, colour = Entity, group = Entity)) +
  geom_line() +
  labs(title = "Trends in Military Expenditures of countries in WW1 \n (from before and after 10 years of WW1)", y = "Expenditure", x = "")

```

Primary School Enrolment Percentage Trends

```{r}
school_enrolment <- subset(primary_school_main, Entity %in% world_war1$StateName)
school_enrolment <- subset(school_enrolment, Year >= min(world_war1$StartYear1)-10)
school_enrolment <- subset(school_enrolment, Year <= max(world_war1$EndYear1)+10)
ggplot(school_enrolment,aes(Year, y = SchoolEnrolment, colour = Entity, group = Entity)) +
  geom_line() +
  labs(title = "Trends in Primary School enrolment in countries in WW1", y = "Enrolment", x = "")
```

Human Development Index Trends
```{r}
HDI_table <- subset(HDI_main, Entity %in% world_war1$StateName)
ggplot(HDI_table,aes(Year, y = HDI, colour = Entity)) +
  geom_line() +
  labs(title = "Trends in Human Development Index in countries in WW1", y = "HDI Value", x = "")

HDI_table <- subset(HDI_table, Year >= min(world_war1$StartYear1)-10)
HDI_table <- subset(HDI_table, Year <= max(world_war1$EndYear1)+10)
ggplot(HDI_table,aes(Year, y = HDI, colour = Entity)) +
  geom_line() +
  labs(title = "Trends in Human Development Index in countries \n 5 years before and after WW1", y = "HDI Value", x = "")

```

World War 2
```{r}
#world_war2:: Years: 1939-1945
world_war2=inter_state%>%filter(WarName=='World War II')
world_war2$StateName=as.character(world_war2$StateName)
#maps
world <- map_data("world")
world_war2 <- world_war2 %>%
  ## Recode certain entries
  mutate(StateName = recode(str_trim(StateName), "United States of America" = "USA",
                            "United Kingdom" = "UK"))

```

Military Expenditure Trends

```{r}
library(ggplot2)
military_expenditure <- subset(military_expenditure_main, Entity %in% world_war2$StateName)
  ggplot(military_expenditure,aes(Year, y = MilitaryExpenditure, colour = Entity, group = Entity)) +
  geom_line() +
  labs(title = "Trends in Military Expenditures of countries in WW2", y = "Expenditure", x = "")
  
  library(ggplot2)
military_expenditure <- subset(military_expenditure_main, Entity %in% world_war2$StateName)
military_expenditure <- subset(military_expenditure, Year >= min(world_war2$StartYear1)-10)
military_expenditure <- subset(military_expenditure, Year <= max(world_war2$EndYear1)+10)
 ggplot(military_expenditure,aes(Year, y = MilitaryExpenditure, colour = Entity, group = Entity)) +
  geom_line() +
  labs(title = "Trends in Military Expenditures of countries in WW2 \n (from before and after 10 years of WW2)", y = "Expenditure", x = "")

```

Primary School Enrolment Percentage Trends

```{r}
school_enrolment <- subset(primary_school_main, Entity %in% world_war2$StateName)
school_enrolment <- subset(school_enrolment, Year >= min(world_war2$StartYear1)-10)
school_enrolment <- subset(school_enrolment, Year <= max(world_war2$EndYear1)+10)
ggplot(school_enrolment,aes(Year, y = SchoolEnrolment, colour = Entity, group = Entity)) +
  geom_line() +
  labs(title = "Trends in Primary School enrolment in countries in WW2", y = "Enrolment", x = "")
```

Human Development Index Trends
```{r}
HDI_table <- subset(HDI_main, Entity %in% world_war2$StateName)
ggplot(HDI_table,aes(Year, y = HDI, colour = Entity)) +
  geom_line() +
  labs(title = "Trends in Human Development Index in countries in WW2", y = "HDI Value", x = "")

HDI_table <- subset(HDI_table, Year >= min(world_war2$StartYear1)-10)
HDI_table <- subset(HDI_table, Year <= max(world_war2$EndYear1)+10)
ggplot(HDI_table,aes(Year, y = HDI, colour = Entity)) +
  geom_line() +
  labs(title = "Trends in Human Development Index in countries \n 5 years before and after WW2", y = "HDI Value", x = "")

```

