---
title: "EDAV Fall 2019 PSet5 Analysis"
output: html_document
---


```{r, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE)

# install and load the libraries
if(!require("ggplot2")) { install.packages("ggplot2"); library("ggplot2")}
if(!require("dplyr")) { install.packages("dplyr"); library("dplyr")}
if(!require("tidyverse")) { install.packages("tidyverse"); library("tidyverse")}
if(!require("maps")) { install.packages("maps"); library("maps")}
if(!require("GGally")) { install.packages("GGally"); library("GGally")}
if(!require("RColorBrewer")) { install.packages("RColorBrewer"); library("RColorBrewer")}
if(!require("parcoords")) { install.packages("parcoords"); library("parcoords")}
if(!require("gridExtra")) { install.packages("gridExtra"); library("gridExtra")}
if(!require("cowplot")) { install.packages("cowplot"); library("cowplot")}
```


### 2. The Questions


List three questions that you hope you will be able to answer from your research.

* General visualization explorations of wars, such as:
    + what types of wars happened most frequently (done),
    + which country initiated inter-state wars most frequently, 
    + which war type caused the most number of deaths,  
    + what wars caused the most number of deaths, 
    + is the number of deaths correlated to whether the country is an initiator of an inter-state war or not, 
    + is there a pattern between initiators and winners of inter-state wars, 
    + what countries had most losses of human lives due to the war, 
    + what countries participated most actively in the wars (done), 
    + is there a pattern of the starting month of all wars, etc.

* What are the similarities and differences between World War I and World War II? (e.g., How were the global food prices affected during WW1 and WW2 time periods?)

* What are the effects of wars on the economics, food production, population, and education of some countries? (We may focus on some countries like Japan, Russia, Germany, China, United Kingdom, France that both participated in WWI and WWII.)  



```{r}
# import data
inter_state <- read.csv('data/raw/Inter-StateWarData_v4.0.csv')
intra_state <- read.csv('data/raw/Intra-StateWarData_v4.1.csv')
extra_state <- read.csv('data/raw/Extra-StateWarData_v4.0.csv')
non_state <- read.csv('data/raw/Non-StateWarData_v4.0.csv')
```

Intra state wars are further categorized into 3 types as follows:


```{r}
Civil_intra_state <- intra_state %>% filter(WarType==4 | WarType==5)
Regional_internal_intra_state <- intra_state %>% filter(WarType==6)
Intercommunal_intra_state <-intra_state %>% filter(WarType==7)
```


#### a.1 what types of wars happened most frequently

```{r}
intra <- intra_state %>% 
  select(WarNum, WarName, WarType) %>% 
  mutate(Type="Intra-state") %>%
  mutate(SubType=ifelse(WarType==4, "4: Intra-Civil war for central control",
                        ifelse(WarType==5, "5: Intra-Civil war over local issues",
                               ifelse(WarType==6, "6: Intra-Regional internal","7: Intra-Intercommunal"))))

inter <- inter_state %>% 
  select(WarNum, WarName, WarType) %>% 
  mutate(Type="Inter-state") %>%
  mutate(SubType="1: Inter")

extra <- extra_state %>% 
  select(WarNum, WarName, WarType) %>% 
  mutate(Type="Extra-state") %>%
  mutate(SubType=ifelse(WarType==2, "2: Extra-Colonial", "3: Extra-Imperial"))

non <- non_state %>% 
  select(WarNum, WarName, WarType) %>% 
  mutate(Type="Non-state") %>%
  mutate(SubType=ifelse(WarType==8, "8: Non-In nonstate territory", "9: Non-Across state borders"))

total <- rbind(intra, inter, extra, non)
total$Type<- as.factor(total$Type)
total$SubType<- as.factor(total$SubType)
```



```{r}
total_stacked <- total %>%
  group_by(Type, SubType) %>%
  summarise(count = n())

# http://www.sthda.com/english/wiki/wiki.php?id_contents=7930#add-a-common-legend-for-multiple-ggplot2-graphs
get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

p1<- ggplot(total_stacked, aes(y=count, x=Type, fill=SubType)) + 
  geom_bar(stat="identity") + 
  ggtitle("Figure 1: Stacked bar chart\non war types") + 
  theme(plot.title = element_text(hjust = 0.5, size=12))

p2 <- ggplot(total_stacked, aes(y=count, x=Type, fill=SubType)) + 
  geom_bar(position="fill", stat="identity") + 
  ylab("percent") +
  ggtitle("Figure 2: Percent stacked bar chart\non war types") + 
  theme(plot.title = element_text(hjust = 0.5, size=12),
        legend.position = "none") +
  scale_y_continuous(labels = scales::percent_format()) 

legend <- get_legend(p1)

p1 <- p1 + theme(legend.position="none")

lankPlot <- ggplot() + 
  geom_blank(aes(1,1)) + 
  cowplot::theme_nothing()
```



```{r fig.width=10}
grid.arrange(p1, p2, legend,
             ncol=3, nrow = 1, 
             widths = c(2.5,2.5,1.5), 
             heights = c(4))
```






#### b. similarities and differences between World War I and World War II


```{r}
df1=inter_state%>%group_by(StateName)%>%summarise(total_deaths=sum(BatDeath))

world <- map_data("world")

df1 <- df1 %>%
  ## Recode certain entries
  mutate(StateName = recode(str_trim(StateName), 
                            "United States of America" = "USA",
                            "United Kingdom" = "UK",
                            "China (PRC)"="China",
                            "Taiwan (ROC)"="Taiwan",
                            "Yemen Arab Republic"="Yemen"))

df1$total_deaths=df1$total_deaths/1000

ggplot(df1, aes(map_id = StateName)) + 
  geom_map(aes(fill = total_deaths), map = world) +
  expand_limits(x = world$long, y = world$lat) +
  borders(database = "world", 
          regions = ".", 
          fill = NA,
          colour = "grey50", 
          xlim = NULL, 
          ylim = NULL) +
  scale_fill_distiller(type = "seq", 
                       palette = "Spectral",
                       direction = -1,
                       limits = c(min(df1$total_deaths), max(df1$total_deaths)), 
                       na.value = "white", 
                       name = "Number of deaths in thousands") +
  coord_fixed(1.3) +
  labs(title = "Figure 2: Deaths of Inter-state Wars by Countries (1816-2007)", x=" ", y=" ")
```




```{r}
# maps
world <- map_data("world") %>%
  mutate(region = recode(str_trim(region), 
                            "Russia" = "USSR",
                            "Georgia" = "USSR",
                            "Ukraine" = "USSR", 
                            "Moldova" = "USSR", 
                            "Belarus" = "USSR", 
                            "Armenia" = "USSR", 
                            "Azerbaijan" = "USSR", 
                            "Kazakhstan" = "USSR", 
                            "Uzbekistan" = "USSR", 
                            "Turkmenistan" = "USSR", 
                            "Kyrgyzstan" = "USSR", 
                            "Tajikistan" = "USSR",
                         "Bosnia and Herzegovina" = "Yugoslavia",
                         "Croatia" = "Yugoslavia",
                         "Macedonia" = "Yugoslavia",
                         "Montenegro" = "Yugoslavia",
                         "Serbia" = "Yugoslavia",
                         "Slovenia" = "Yugoslavia", 
                         "Czech Republic" = "Austria-Hungary",
                         "Hungary" = "Austria-Hungary",
                         "Montenegro" = "Austria-Hungary",
                         "Poland" = "Austria-Hungary",
                         "Serbia" = "Austria-Hungary",
                         "Slovakia" = "Austria-Hungary",
                         "Slovenia" = "Austria-Hungary",
                         "Ukraine" = "Austria-Hungary"))


Russia <- map_data("world") %>% filter(region == "Russia")
world <- rbind(world, Russia)
```



```{r}
# world_war1:: Years: 1914-1918
world_war1=inter_state%>%filter(WarName=='World War I')
world_war1$StateName=as.character(world_war1$StateName)

world_war1 <- world_war1 %>%
  ## Recode certain entries
  mutate(StateName = recode(str_trim(StateName), 
                            "United States of America" = "USA",
                            "United Kingdom" = "UK"))

world_war1 <- world_war1 %>% mutate(number_of_days = as.Date(with(world_war1, paste(EndYear1, EndMonth1, EndDay1,sep="-")), "%Y-%m-%d")-as.Date(with(world_war1, paste(StartYear1, StartMonth1, StartDay1,sep="-")), "%Y-%m-%d"))

world_war1$number_of_days <- as.numeric(world_war1$number_of_days )

# world_war2:: Years: 1939-1945
world_war2=inter_state%>%filter(WarName=='World War II')
world_war2$StateName=as.character(world_war2$StateName)

# maps
world_war2 <- world_war2 %>%
  ## Recode certain entries
  mutate(StateName = recode(str_trim(StateName), 
                            "United States of America" = "USA",
                            "United Kingdom" = "UK"))


world_war21 <- world_war2 %>% filter(StartYear2 !=-8) 

world_war21=world_war21%>%  mutate(number_of_days = as.Date(with(world_war21, paste(EndYear1, EndMonth1, EndDay1,sep="-")), "%Y-%m-%d")-as.Date(with(world_war21, paste(StartYear1, StartMonth1, StartDay1,sep="-")), "%Y-%m-%d")+as.Date(with(world_war21, paste(EndYear2, EndMonth2, EndDay2,sep="-")), "%Y-%m-%d")-as.Date(with(world_war21, paste(StartYear2, StartMonth2, StartDay2,sep="-")), "%Y-%m-%d"))

world_war22 <- world_war2 %>% filter(StartYear2 ==-8)

world_war22=world_war22 %>%  mutate(number_of_days = as.Date(with(world_war22, paste(EndYear1, EndMonth1, EndDay1,sep="-")), "%Y-%m-%d")-as.Date(with(world_war22, paste(StartYear1, StartMonth1, StartDay1,sep="-")), "%Y-%m-%d"))

world_war2=rbind(world_war21,world_war22)

world_war2$number_of_days <- as.numeric(world_war2$number_of_days)


world_wars <- rbind(world_war1, world_war2)

world_wars$WarName <- as.factor(world_wars$WarName)

write_csv(world_wars, "world_wars.csv")
```

```{r}
world2 <- map_data("world") %>% filter(region == "Bosnia and Herzegovina")
for (val in unique(world_war1$StateName) ) {
  if (!(val %in%  unique(world$region)))  {print(val)}}
```




```{r fig.width=12, fig.height=5}
ggplot(world_wars, aes(map_id = StateName)) + 
  geom_map(aes(fill = BatDeath), map = world, col = "white") +
  expand_limits(x = world$long, y = world$lat) +
  scale_fill_distiller(name = "Number of Battle Deaths", palette = "Spectral") +
  borders(database = "world", 
          regions = ".", 
          fill = NA,
          colour = "grey50", 
          xlim = NULL, 
          ylim = NULL) +
  labs(title = "Battle Deaths for World War I versus World War II", 
       x = "Longitude", 
       y = "Latitude") +
  facet_wrap(~WarName, scales = "free", ncol = 2) +
  theme(plot.title = element_text(hjust = 0.5, size = 16),
        strip.text.x = element_text(size = 8),
        legend.position="top",
        legend.key.width=unit(1.5,"cm"))
```




```{r fig.width=12, fig.height=5}
ggplot(world_wars, aes(map_id = StateName)) + 
  geom_map(aes(fill = number_of_days), map = world, col = "white") +
  expand_limits(x = world$long, y = world$lat) +
  scale_fill_distiller(name = "Number of Battle Deaths", palette = "Spectral") +
  borders(database = "world", 
          regions = ".", 
          fill = NA,
          colour = "grey50", 
          xlim = NULL, 
          ylim = NULL) +
  labs(title = "Number of Days for World War I versus World War II", 
       x = "Longitude", 
       y = "Latitude") +
  facet_wrap(~WarName, scales = "free", ncol = 2) +
  theme(plot.title = element_text(hjust = 0.5, size = 16),
        strip.text.x = element_text(size = 8),
        legend.position="top",
        legend.key.width=unit(1.5,"cm"))
```
```



